---
title: "Assignment5_VoiceInASD"
author: "Riccardo Fusaroli"
date: "July 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 5 - Assessing voice in ASD

People with ASD have been often described as odd in their speech. A recent meta-analysis of the literature (Fusaroli et al. 2016) indicates that pitch might be involved.

In the course of this assignment we will:
- extract parametric and non-parametric descriptors of pitch
- extract dynamic descriptors of pitch
- compare how well these descriptors characterize the autistic voice *relative to each other*
- assess whether these descriptors complement each other in describing the autistic voice


## Load libraries and data
```{r}


```

## Extract parametric descriptors

Mean, standard deviation, range

```{r}

```

## Extract non-parametric descriptors

Median, interquartile range, coefficient of variation, mean absoluted deviation, median absolute deviation, skewness, kurtosis

```{r}

```

## Extract dynamic descriptors

Recurrence Quantification Analysis, Power law alpha, DFA alpha

```{r}

```

## Let's figure out which single descriptor best discriminates

N.B. remember the first two classes: cross-validate or use information criteria :-)

```{r}

```

## Let's explore whether the descriptors complement each other (find the best combination of descriptors discriminating the voice of ASD from the voice of TD)

N.B. remember the first two classes: cross-validate or use information criteria :-)

```{r}

```

## Report the results

METHODS SECTION: how did you analyse the data?

RESULTS SECTION: which descriptor is the single best descriptor? what's its accuracy, sensitivity, specificity? what's the best combination of 