---
title: "Assignment 1 - Language Development in ASD"
author: "Riccardo Fusaroli"
date: "July 7, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Language development in Autism Spectrum Disorder (ASD)

Background: Autism Spectrum Disorder is often related to language impairment. However, this phenomenon has not been empirically traced in detail: i) relying on actual naturalistic language production, ii) over extended periods of time. We therefore videotaped circa 30 kids with ASD and circa 30 comparison kids (matched by linguistic performance at visit 1) for ca. 30 minutes of naturalistic interactions with a parent. We repeated the data collection 6 times per kid, with 4 months between each visit. We transcribed the data and counted: 
i) how many words does the kid use in each video? How many does the parent? 
ii) how many unique words does the kid use in each video? How many does the parent? 
iii) how many morphemes per utterance (Mean Length of Utterance) does the child use? And the parent? 
This data is in the file "Assignment1_LanguageDevelopmentInASD.csv". NB. Not all children are present there. We will use the missing ones next week.
We then want to test the language trajectory of child and parent over time.

This Markdown is structured in the following way:

1. The exercises: read them carefully. Under each exercise you will have to write your answers, once you have written and run the code.
2. A guided template full of hints for writing the code to solve the exercises. Fill in the code and the paragraphs as required. Then report your results under the exercise part.
3. In exercise 4 you will be asked to create the best possible model of language development in TD and ASD children, picking and choosing whatever additional variables you want from the dataset. Next time, the models produced by the different groups will compete against each other to see who can produce the best model, so choose carefully!

## Exercise 1) Preliminary Data Exploration
Describe the participant samples in the dataset. Do you think the two groups are well balanced? If not, what do you think was the reason?

[HERE GOES YOUR ANSWER]

### Exercise 2) Children learning language: the effects of time and ASD
Describe linguistic development in TD and ASD children. Does it vary across the three different measures of linguistic performance (total words, unique words, MLU)?

[HERE GOES YOUR ANSWER]

### Exercise 3) Child directed speech as a moving target
Describe how parental use of language changes over time. What do you think is going on?

[HERE GOES YOUR ANSWER]

### Exercise 4) Looking into "individual differences" (demographic, clinical or cognitive profiles)
The dataset contains some additional variables characterizing the kids’ cognitive and clinical profile: ADOS (autism severity), MSEL EL (Expressive Language, that is, linguistic skills at first visit as assessed by a psychologist using Mullen Scales of Early Learning), MSEL VR (Visual Reception, used as a proxy for non verbal IQ at first visit), Age, Gender, Ethnicity. Would it make sense to add any of them to your model of linguistic trajectories? Create the best possible model (the one that best explain the data. One for MLU, one for unique words, one for total words). Next time your model will be tested on new participants, and we will proclaim a winner. Describe your strategy to select the best models (how did you choose the variables to include?) and send the code to Riccardo by next tuesday.

[HERE GOES YOUR ANSWER]

### [OPTIONAL] Exercise 5 ) Comment on how the three linguistic variables measure linguistic performance (the so-called "construct validity" of the measures). Do they express the same variance?

[HERE GOES YOUR ANSWER]

### Structure of the code chunks

Basic stuff:
- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there? Are they in the right format?) and describe the participants (by diagnosis)

We will try to answer three questions:

- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
- Which variables should we use to best explain the child linguistic performance?
  
### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
- e.g. mixed effects models

```{r Load Libraries}

```

### Define your working directory and load the data

- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data: "Assignment1_LanguageDevelopmentInASD.csv" (use read.csv(fileName))

```{r Load Data}

#getwd()
#locpath=
#setwd(locpath)
#Data =

```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, MSEL EL, MSEL VR, Visit, Number of words used, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

```{r}

```

[REPORT THE RESULTS]

## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis 1a: Number of words used by the child change: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?
- Which variable(s) should be included as random factors?

```{r}

```

How would you evaluate whether the model is a good model?

```{r}

```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better

```{r}

```

Exciting right?
Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results

[REPORT THE RESULTS]
Linguistic development of total words used in children is affected by ... [COMPLETE]

Let's now apply what we have learned to the other indexes of child language performance and parent language performance.

### Hypothesis 1b: Number of unique words used by the child change: i) over time, ii) according to diagnosis

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors?
- Which variable(s) should be included as random factors?

```{r}

```

[REPORT THE RESULTS]

### Hypothesis 1c: Length of utterance used by the child changes: i) over time, ii) according to diagnosis

```{r}

```

[REPORT THE RESULTS]

## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis 2a: number of words used by the parent change: i) over time, ii) according to diagnosis

Remember to plot the data first and then to run a statistical test.

```{r}

```

[REPORT THE RESULTS]

### Hypothesis 2b: Unique words used by the parent change: i) over time, ii) according to diagnosis

Remember to plot the data first and then to run a statistical test.

```{r}

```

[REPORT THE RESULTS]

### Hypothesis 2c: Length of utterance used by the parent changes: i) over time, ii) according to diagnosis

```{r}

```

[REPORT THE RESULTS]

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables. Try them out and identify the statistical models that best describes your data (children's MLU, unique words and total words). Describe how you selected the best model and send the code to run the model to Riccardo by next tuesday.


```{r}

```

REPORT OF THE RESULTS