---
title: "Assignment2_LanguageDevelopmentInASD"
author: "Riccardo Fusaroli"
date: "July 10, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the second exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time, then learning how to cross-validate models and finally how to systematically compare models.

Exercise 1
Exercise 2
Exercise 3


## Train a model and test it - Vanilla Version

```{r}
library(lme4)

model = lmer(ChildMLU~Visit*Diagnosis+(1|Child.ID),trainData,REML=F)
testData2$predict=predict(model,testData2, allow.new.levels = TRUE)
ggplot(testData2,aes(predict,ChildMLU))+geom_point(aes(colour=as.factor(Child.ID)))
model1=lmer(ChildMLU~predict+(1|Child.ID),testData2,REML=F)
r.squaredGLMM(model1)

```

## Train a model and test it - Cross-Validation

```{r}

```