---
title: "Assignment4_PowerLaws"
author: "Riccardo Fusaroli"
date: "July 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 4 - Assessing Complex Systems via Power Laws

In this assignment you will analyse the finger tapping data you produced last week

- let's visualize the distribution
- let's check for normality
- let's check for time-dependence (auto-correlation, mi). But this are local patterns!

- Let's check for power-law distribution!
- Let's check for spectral density (noise analysis)

- Let's loop through all the time-series
- Let's run a mixed effects model to figure out whether:
  a) there is a higher amount of powerlaws in one condition
  b) the alpha estimation is different between conditions

N.B. Should this be first with a simulated time-series? Or straight to the fingertapping?

```{r}
# Generating white noise for comparison
values=rnorm(1000, mean = 0, sd = 1)
wNoise=data.frame(values)
# Generating random numbers from a power law
values=scale(rpldis(1000, xmin=1, alpha=1.1, discrete_max = 10000),center=T,scale=T)
pLaw1=data.frame(values)
values=scale(rpldis(1000, xmin=1, alpha=2, discrete_max = 10000),center=T,scale=T)
pLaw2=data.frame(values)
values=scale(rpldis(1000, xmin=1, alpha=3, discrete_max = 10000),center=T,scale=T)
pLaw3=data.frame(values)

```

## Visualizing the distributions

```{r}
ggplot(wNoise,aes(values))+geom_histogram()
ggplot(pLaw1,aes(values))+geom_histogram()
ggplot(pLaw2,aes(values))+geom_histogram()
ggplot(pLaw3,aes(values))+geom_histogram()
```

## Check for normality

```{r}
shapiro.test(wNoise$values)
qqnorm(wNoise$values);qqline(wNoise$values, col = 2)
shapiro.test(pLaw1$values)
qqnorm(pLaw1$values);qqline(pLaw1$values, col = 2)
shapiro.test(pLaw2$values)
qqnorm(pLaw2$values);qqline(pLaw2$values, col = 2)
shapiro.test(pLaw3$values)
qqnorm(pLaw3$values);qqline(pLaw3$values, col = 2)
```


## Check for time-dependence

```{r}
tau.ami = timeLag(pLaw2$values, technique = "ami", lag.max = 30, do.plot = T)
emb.dim = estimateEmbeddingDim(pLaw2$values, time.lag = tau.ami,max.embedding.dim = 10)

```

## Check for power-law distribution!

N.B. this follows step-by-step the tutorial of poweRlaw

```{r}

# Create a new power law object
m_m = conpl$new(pLaw2$values)
m_m$getXmin()
m_m$getPars()

#m_m$setPars(2)
# Estimate xmin (using kolmogorov-smirnoff)
(est = estimate_xmin(m_m))
# Set xmin
m_m$setXmin(est)
# Estimate alpha given an xmin
(est = estimate_pars(m_m))

# Create the data to plot the powerlaw
dd = plot(m_m)
ggplot(dd, aes(x,y))+geom_point() + scale_y_log10()+ scale_x_log10()

# Now estimating uncertainty
cores=parallel::detectCores() # Figuring out how many cores on your machine
bs = bootstrap(m_m, no_of_sims=1000, threads=cores) # Bootstrapping and distributing the bootstrapping across cores
hist(bs$bootstraps[,2], breaks="fd")#xmin
hist(bs$bootstraps[,3], breaks="fd")#alpha
plot(jitter(bs$bootstraps[,2], factor=1.2), bs$bootstraps[,3])
bs_p = bootstrap_p(m_m, no_of_sims=1000, threads=cores)
```

## Check for spectral density (noise analysis)

```{r}
x1=dfa(wNoise$values)
estimate(x1,do.plot=TRUE)
#x2=rqa(takens = NULL, time.series = pLaw2$values, embedding.dim = emb.dim, time.lag = tau.ami,radius=1,do.plot = T)
```

## Now let's loop the whole process through the whole set of time-series

Remember that we are interested in having a dataframe with the following columns:
- Participant
- Condition
- Normality
- Lag at which autocorrelation is minimized
- Lag at which mutual information is minimized
- Embedding Dimensions
- Estimated xmin
- Estimated alpha
- ? Something on uncertainty??
- Dfa estimate

```{r}

```

## Now let's test a few hypotheses
Per each hypothesis run a statistical test and a draw a plot

- H1: The metronome condition will have more normal distributions of interbeat intervals than the self-driven one
- H2: 

```{r}

```


